---
title: "Epigenetic Aging Analysis Report"
author: "BIOF3001 Project"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(ggplot2)
library(dplyr)
library(knitr)
```

# Introduction

This report presents the analysis of DNA methylation data to investigate epigenetic aging and age acceleration.

## Background

Epigenetic clocks use DNA methylation patterns to estimate biological age. Age acceleration occurs when epigenetic age exceeds chronological age, potentially indicating accelerated aging processes.

# Data Overview

```{r data-loading, eval=FALSE}
# Load processed data
# data <- readRDS("data/processed/methylation_processed.rds")

# Display summary statistics
# summary(data)
```

## Sample Characteristics

```{r sample-stats, eval=FALSE}
# Display sample information
# table(data$sex)
# summary(data$age)
```

# Epigenetic Age Calculation

## Horvath Clock

```{r horvath-clock, eval=FALSE}
# Calculate Horvath epigenetic age
# epigenetic_age <- calculate_horvath_age(methylation_data)
```

## Age Acceleration

```{r age-acceleration, eval=FALSE}
# Calculate age acceleration
# age_accel <- epigenetic_age - chronological_age
```

# Results

## Age Correlation

```{r age-correlation, eval=FALSE}
# Scatter plot of epigenetic age vs chronological age
# ggplot(data, aes(x = chronological_age, y = epigenetic_age)) +
#   geom_point(alpha = 0.6) +
#   geom_smooth(method = "lm", color = "blue") +
#   geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
#   labs(
#     title = "Epigenetic Age vs Chronological Age",
#     x = "Chronological Age (years)",
#     y = "Epigenetic Age (years)"
#   ) +
#   theme_minimal()
```

## Age Acceleration Distribution

```{r age-accel-dist, eval=FALSE}
# Histogram of age acceleration
# ggplot(data, aes(x = age_acceleration)) +
#   geom_histogram(bins = 30, fill = "steelblue", alpha = 0.7) +
#   geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
#   labs(
#     title = "Distribution of Age Acceleration",
#     x = "Age Acceleration (years)",
#     y = "Count"
#   ) +
#   theme_minimal()
```

## Statistical Tests

```{r statistical-tests, eval=FALSE}
# Correlation test
# cor.test(data$chronological_age, data$epigenetic_age)

# Linear regression
# model <- lm(epigenetic_age ~ chronological_age, data = data)
# summary(model)
```

# Discussion

## Key Findings

1. Relationship between epigenetic and chronological age
2. Factors associated with age acceleration
3. Biological implications

## Limitations

- Sample size considerations
- Platform-specific effects
- Confounding factors

# Conclusions

Summary of main results and their implications for understanding epigenetic aging.

# Session Information

```{r session-info}
sessionInfo()
```
